---
title: Global Rice Production Analysis

# set to 1 to the authors and acknowledgements, producing the required blinded version.
blinded: 1

authors: 
- name: Hari Prasannaa Thangavel Ravi
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of Data Science, George Washington University
  
- name: Jin Hyuk Son
  affiliation: Department of Data Science, George Washington University

- name: Pranjal S. Wakpaijan
  affiliation: Department of Data Science, George Washington University

keywords:
- 3 to 6 keywords
- that do not appear in the title
abstract:
  Predicting rice production outcomes is a critical aspect of global food security and agricultural
  sustainability. Understanding the factors that influence rice yield helps farmers and policymakers
  make informed decisions, ultimately improving food supply and nutrition worldwide. This project,
  titled "Global Rice Production Analysis," utilizes a dataset named “Explore Data on Agricultural
  Production,” which contains approximately 13,500 rows and roughly 40 columns. We aim to
  explore the intricate dynamics of rice production by analyzing key parameters such as land area
  used for cultivation, harvested area, climatic conditions, and factors related to waste generation.
  We will begin with exploratory data analysis (EDA) to clean the dataset, addressing missing
  values and outliers while examining data distributions and inter-relationships among variables.
  By evaluating various factors such as soil quality, rainfall patterns, and agricultural practices we
  aim to build predictive models that can estimate rice yields. This study not only addresses
  real-world challenges in rice cultivation but also provides valuable insights to enhance
  agricultural strategies, optimize resource allocation, and promote sustainable practices. The
  findings will support rice producers in improving yield predictions and ultimately contribute to
  enhancing global food security.

bibliography: bibliography.bib
biblio-style: apalike # default is using plain.bst
output: rticles::asa_article
---

# Introduction

This template demonstrates some of the basic latex you'll need to know to create a ASA article.
It is built from template find at Journal of the American Statistical Association [instruction for authors](https://www.tandfonline.com/action/authorSubmission?show=instructions&journalCode=uasa20#formatting)

## Plot

Plot example - LaTeX floating to be adjusted if you don't want it.

```{r, fig.cap = "A plot example", echo = FALSE}
plot(cars)
```

## Table

Table example

```{r, echo = FALSE}
knitr::kable(head(cars))
```

## Various guidelines

- Note that figures and tables should appear in the paper, not at the end or
in separate files.
- Remember that in the blind version, you should not identify authors indirectly in the text.  That is, don't say "In Smith et. al.  (2009) we showed that ...".  Instead, say "Smith et. al. (2009) showed that ...".
- These points are only intended to remind you of some requirements. Please refer to the instructions for authors
at <http://amstat.tandfonline.com/action/authorSubmission?journalCode=uasa20&page=instructions#.VFkk7fnF_0c>
- For more about ASA style, please see <http://journals.taylorandfrancis.com/amstat/asa-style-guide/>
- If you have supplementary material (e.g., software, data, technical proofs), identify them in the section below.  In early stages of the submission process, you may be unsure what to include as supplementary material.  Don't worry---this is something that can be worked out at later stages.

# Methods {#meth}

Don't take any of these section titles seriously.  They're just for illustration.

# Verifications {#verify}

This section will be just long enough to illustrate what a full page of
text looks like, for margins and spacing.

```{=latex}
\addtolength{\textheight}{.5in}%
```

See @Campbell02 and @Schubert13 or [@Chi81]

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
**With this spacing we have 25 lines per page.**
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

```{=latex}
\addtolength{\textheight}{-.3in}%
```

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.

# Conclusion {#conc}

\bigskip
::: {.center latex=true}
**SUPPLEMENTARY MATERIAL**
:::

> Use [definition list syntax](https://pandoc.org/MANUAL.html#definition-lists) in this part

Title:
  ~ Brief description. (file type)

R-package for  MYNEW routine:
  ~ R-package **MYNEW** containing code to perform the diagnostic methods described in the article. The package also contains all datasets used as examples in the article. (GNU zipped tar file)

HIV data set:
  ~ Data set used in the illustration of MYNEW method in Section~ 3.2. (.txt file)

# Bibliography. 

Using `natbib` is the default with this format, using `plain.bst` by default on the template, and `apalike.bst` in this Rmd skeleton. Chante to your preference using the `biblio-style` in YAML


